---
title: "Preliminary Exploratory Analysis of DI and SPI/SDG Relationship"
author: "Sevastian Sanchez"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: no
    toc_depth: '3'
    number_sections: yes
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: yes
    number_sections: yes
    highlight: tango
    theme: default
    fig_caption: yes
    df_print: tibble
---


### NAIVE OLS Models: DI variable on SPI & SDG Performance 
Finding estimated impact of DI on SPI and SDG status prior to adding controls or robust SEs
```{r}
# 1. OLS for SPI and DI - Overall
ols_spi_di_naive <- lm(spi_comp ~ di_score, data = merged)
summary(ols_spi_di_naive)

# 2. OLS for SDG and DI - Overall
ols_sdg_di_naive <- lm(sdg_overall ~ di_score, data = merged)
summary(ols_sdg_di_naive)

# stargazer summary table of models: ols_spi_di_naive, ols_sdg_di_naive.
stargazer(
  ols_spi_di_naive, 
  ols_sdg_di_naive, 
  title = "Naive OLS Models: DI & SPI/SDG",
  dep.var.caption = "Dependent Variable:",
  dep.var.labels = c("SDG Composite Score", "SPI Composite Score"),
  #covariate.labels = c("di_score", "Intercept"),
  #column.labels = c("SPI Only", "SCI Only", "SPI + SCI"),  # <-- Model labels at the top
  omit.stat = c("f", "ser"), 
  digits = 4,
  type = "text"
  #out = "component_1/figures/naive_mods_sdgs&spi_di_tab.html"  # Save as HTML file
)

```
*results*
ols_spi_di_naive: 5.0122 (p-value < 0.001)
ols_sdg_di_naive: 3.22301 (p-value < 0.001)

The naive OLS models indicate that the Democracy Index (DI) has a strong positive relationship with both the Statistical Performance Index (SPI) and the Sustainable Development Goals (SDG) composite scores. The coefficients suggest that a one-unit increase in DI is associated with a 5.0122 increase in SPI and a 3.22301 increase in SDG scores, both statistically significant at p < 0.001. This suggests that countries with higher democracy scores tend to have better statistical performance and SDG outcomes.

### Checking for Heteroskedasticity: residual plots [no need to report]
```{r}
plot(ols_spi_di_naive, which = 1)  # SPI ~ DI model 
plot(ols_sdg_di_naive, which = 1)  # SDG ~ DI model 
```

