---
title: "Interactions & Subgroup Analysis (POLS) [EDITING STAGES]"
author: "Sevastian Sanchez"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: no
    toc_depth: '3'
    number_sections: yes
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: yes
    number_sections: yes
    highlight: tango
    theme: default
    fig_caption: yes
    df_print: tibble
---
```{r include=FALSE}
knitr::opts_chunk$set(
  tidy = TRUE,
  tidy.opts = list(width.cutoff = 60)  # Adjust width as needed
)
```

### Setup
```{r}
# set working directory
setwd("~/Documents/GitHub/QMSS_Thesis_Sanchez")

# load libraries
source("packages.R")

#load final merged df 
merged <- read_csv("data/Main CSV Outputs/comp1_data.csv") # updated 

# refer to 'wrangled/adjust_output.Rmd' to make necessary adjustments 
```

### Selecting variables and filtering years +2015
```{r}
# Wrangling for analysis and multicollinearity concerns 
merged_2015 <- merged %>% 
  filter(year >= 2016) %>% 
  # Selecting relevant variables for analysis 
  dplyr::select(country_code, year, sdg_overall, spi_comp, di_score, log_gdppc, income_level, regime_type_2, regime_type_4, regch_event)
```

### Test 3: Checking for Interactions
modeling how the effect of SPI on SDG performance depends on one/any of the following moderators:
  - income level (GNI Classification)
  - regime type (RoW Binary & Categorical)
  - democracy index (di_score)
- Is there a need for subgroup analysis, and if so, by what kind of group? 
- Possible Moderators: GNI Classification (income_level), regime_type_2, regime_type_4, di_score

#### Recoding for Interactions
```{r}
# Recode to dummy variables
merged_2015 <- merged_2015 %>% 
  # Recoding income_level, split income_level into dummy variables using case_when()
  mutate(income_level_recoded = case_when(
    income_level == "L" ~ 0, # Low-Income
    income_level == "LM" ~ 1, # Lower-Middle-Income
    income_level == "UM" ~ 2, # Upper-Middle-Income
    income_level == "H" ~ 3, # High-Income
    TRUE ~ NA_real_  # Handle any other cases
  )) %>% 
  mutate(income_level_recoded = as.factor(income_level_recoded)) %>% 
  # recoding/factorizing regime_type_2: 0 = Autocracy; 1 = Democracy 
  mutate(regime_type_binary = as.factor(regime_type_2)) %>% 
  # recoding regime_type_4: 0 = Closed Autocracy; 1 = Electoral Autocracy; 2 = Electoral Democracy; 3 = Full Democracy
  mutate(regime_type_categ = as.factor(regime_type_4)) %>% 
  mutate(aut_ep = as.factor(aut_ep), # autocratization episode
         dem_ep = as.factor(dem_ep)) # democratization episode

### HYPOTHESES ###
# H0: NUll, the impact of SPI on SDG Performance does NOT vary depending on Z [income_level_recoded, regime_type_binary, regime_type_categ, di_score]
# H1: The impact of SPI on SDG Performance varies depending on Z [income_level_recoded, regime_type_binary, regime_type_categ, di_score]

# Interaction 1: 
# HA: The impact of SPI on SDG performance varies depending on GNI Classification (income_level_recoded: 1 = Low; 2 = Lower-Middle; 3 = Upper-Middle; 4 = High)
spi_x_inc_lvl <- plm(sdg_overall ~ cen_spi_comp*factor(income_level_recoded) + cen_di_score +
                       cen_log_gdppc + I(cen_log_gdppc^2) + factor(year), 
                     model = "pooling", 
                     index = c("country_code", "year"),
                     data = merged_2015)
summary(spi_x_inc_lvl, vcov = vcovHC(spi_x_inc_lvl, cluster = "group", type = "HC1"))


# Interaction 2: 
# HA: The impact of SPI on SDG performance varies depending on Regime Type (Binary: 0 = Autocracy; 1 = Democracy)
spi_x_reg_binary <- plm(sdg_overall ~ cen_spi_comp*factor(regime_type_binary) + cen_di_score +
                          cen_log_gdppc + I(cen_log_gdppc^2) + factor(year), 
                        model = "pooling", 
                        index = c("country_code", "year"),
                        data = merged_2015)
summary(spi_x_reg_binary, vcov = vcovHC(spi_x_reg_binary, cluster = "group", type = "HC1"))


# Interaction 3: 
# HA: The impact of SPI on SDG performance varies depending on Regime Type (Categorical: 0 = Closed Autocracy; 1 = Electoral Autocracy; 2 = Electoral Democracy; 3 = Full Democracy)
spi_x_reg_categ <- plm(sdg_overall ~ cen_spi_comp*factor(regime_type_categ) + cen_di_score + 
                         cen_log_gdppc + I(cen_log_gdppc^2) + factor(year), 
                       model = "pooling", 
                       index = c("country_code", "year"),
                       data = merged_2015)
summary(spi_x_reg_categ, vcov = vcovHC(spi_x_reg_categ, cluster = "group", type = "HC1"))


# Interaction 4: 
# HA: The impact of SPI on SDG performance varies depending on Democracy Score (cen_di_score: 0-1, continuous)
spi_x_di_score <- plm(sdg_overall ~ cen_spi_comp*cen_di_score + cen_log_gdppc + I(cen_log_gdppc^2) +
                        factor(year), 
                      model = "pooling", 
                      index = c("country_code", "year"),
                      data = merged_2015)
summary(spi_x_di_score, vcov = vcovHC(spi_x_di_score, cluster = "group", type = "HC1"))

# Interaction 5
# HA: The impact of SPI on SDG performance varies depending on Regime Change Status/Direction
spi_x_reg_eps <- plm(sdg_overall ~ cen_spi_comp*aut_ep + cen_spi_comp*dem_ep +
                     cen_log_gdppc + I(cen_log_gdppc^2) + factor(year), 
                   model = "pooling", 
                   index = c("country_code", "year"),
                   data = merged_2015)
summary(spi_x_reg_eps, vcov = vcovHC(spi_x_reg_eps, cluster = "group", type = "HC1"))

# Final Model 
final_model <- plm(sdg_overall ~ cen_spi_comp*factor(income_level_recoded) + cen_spi_comp*cen_di_score + cen_log_gdppc + I(cen_log_gdppc^2) + factor(year), 
                   model = "within", 
                   index = c("country_code", "year"),
                   data = merged_2015)
summary(final_model, vcov = vcovHC(final_model, cluster = "group", type = "HC1"))
```
*Interaction 1: GNI Income Classification:* 
*HA: The impact of SPI on SDG performance varies depending on GNI Classification (income_level_recoded: 1 = Low; 2 = Lower-Middle; 3 = Upper-Middle; 4 = High)*
[results in doc]...

*Interaction 2: Regime Type (Binary):*
*HA: The impact of SPI on SDG performance varies depending on Regime Type (Binary: 0 = Autocracy; 1 = Democracy)*
...

*Interaction 3: Regime Type (Categorical):*
*HA: The impact of SPI on SDG performance varies depending on Regime Type (Categorical: 0 = Closed Autocracy; 1 = Electoral Autocracy; 2 = Electoral Democracy; 3 = Full Democracy)*
...

*Interaction 4: Democracy Index (Continuous):*
*HA: The impact of SPI on SDG performance varies depending on Democracy Score (cen_di_score: 0-1, continuous)*
...

*Interaction 5: Regime Change Status/Direction:*
*HA: The impact of SPI on SDG performance varies depending on Regime Change Status/Direction*
Results: 


#### Interactions DF 
```{r}
# Create a helper function for tidying plm models with robust SEs
tidy_plm_robust <- function(model, model_name) {
  tidy(model, 
       conf.int = TRUE, 
       vcov. = vcovHC(model, cluster = "group", type = "HC1")) %>%
    mutate(model = model_name,
           n_obs = n_obs)
}

# Enhanced helper function
tidy_plm_robust <- function(model, model_name) {
  tidy_coef <- tidy(model, 
                    conf.int = TRUE, 
                    vcov. = vcovHC(model, cluster = "group", type = "HC1")) %>%
    mutate(model = model_name)
  
  # Extract model statistics
  model_summary <- summary(model)
  tidy_coef %>%
    mutate(
      adj_r_squared = model_summary$r.squared["adjrsq"]
    )
}

# Apply tidy_plm_robust() function to all models
tidy_spi_x_inc_lvl    <- tidy_plm_robust(spi_x_inc_lvl, "Moderator: SPI x Income Level")
tidy_spi_x_reg_binary <- tidy_plm_robust(spi_x_reg_binary, "Moderator: SPI x Regime Binary")
tidy_spi_x_reg_categ  <- tidy_plm_robust(spi_x_reg_categ, "Moderator: SPI x Regime Categ")
tidy_spi_x_di_score   <- tidy_plm_robust(spi_x_di_score, "Moderator: SPI x DI Score")
tidy_spi_x_reg_eps   <- tidy_plm_robust(spi_x_reg_eps, "Moderator: SPI x Regime Change")

model_n_obs <- tibble(
  model = c("Moderator: SPI x Income Level", "Moderator: SPI x Regime Binary", 
            "Moderator: SPI x Regime Categ", "Moderator: SPI x DI Score", "Moderator: SPI x Regime Change"),
  n_obs = c(nobs(spi_x_inc_lvl), nobs(spi_x_reg_binary),
            nobs(spi_x_reg_categ), nobs(spi_x_di_score), nobs(spi_x_reg_eps))
)

# Combine all tidy data frames into one
all_models_tidy <- bind_rows(tidy_spi_x_inc_lvl, 
                             tidy_spi_x_reg_binary, 
                             tidy_spi_x_reg_categ, 
                             tidy_spi_x_di_score, 
                             tidy_spi_x_reg_eps) %>%
  left_join(model_n_obs, by = "model") %>% 
  select(model, term, estimate, std.error, p.value, adj_r_squared, n_obs, everything()) %>%
  arrange(model, term)

# Save the tidy data frame to a CSV file
write_csv(all_models_tidy, "results_csv/interaction_models.csv")


# extracting robust standard errors for each interaction model
rob_se_spi_x_inc_lvl <- coeftest(spi_x_inc_lvl, vcov = vcovHC(spi_x_inc_lvl, cluster = "group", type = "HC1"))
rob_se_spi_x_reg_binary <- coeftest(spi_x_reg_binary, vcov = vcovHC(spi_x_reg_binary, cluster = "group", type = "HC1"))
rob_se_spi_x_reg_categ <- coeftest(spi_x_reg_categ, vcov = vcovHC(spi_x_reg_categ, cluster = "group", type = "HC1"))
#rob_se_spi_x_di_score <- coeftest(spi_x_di_score, vcov = vcovHC(spi_x_di_score, cluster = "group", type = "HC1"))
rob_se_spi_x_reg_eps <- coeftest(spi_x_reg_eps, vcov = vcovHC(spi_x_reg_eps, cluster = "group", type = "HC1"))

# Create a list of robust standard errors for stargazer
robust_se_list <- list(
  rob_se_spi_x_inc_lvl[, 2], 
  rob_se_spi_x_reg_binary[, 2], 
  rob_se_spi_x_reg_categ[, 2], 
  #rob_se_spi_x_di_score[, 2] #,
  rob_se_spi_x_reg_eps[, 2]
  )
          
stargazer(
  spi_x_inc_lvl, spi_x_reg_binary, spi_x_reg_categ, #spi_x_di_score,
  spi_x_reg_eps,
  se = robust_se_list,
  type = "html",
  out = "component_1/figures/pooled_ols_interactions.html",
  title = "Pooled OLS Interaction Models",
  align = TRUE,
  column.labels = c("Income Level", "Regime Binary", "Regime Categ", "Regime Change Episode"),
  dep.var.labels = "SDG Overall",
  omit = "factor\\(year\\)",  # Omits year fixed effects from display
  model.names = FALSE
)

```


## TEST 3b: WB GNI Classifications: income_level ("H", "UM", "LM", "L") 
#### Disaggregated/Grouped by Development Status:
Make 4 regression models and then put them all together in a table to compare the slopes and R-sq values.
```{r}
# 1. Overall model (all countries)
overall_lm <- lm(sdg_overall ~ spi_comp + di_score + log_gdppc + population + factor(year),
                        data = merged_2015)
#summary(overall_lm)
coeftest(overall_lm, vcov = vcovHC(overall_lm, type = "HC1")) #Robust SE

# 2. High income countries
high_inc_lm <- lm(sdg_overall ~ spi_comp + di_score + log_gdppc + population + factor(year),
                        data = merged_2015 %>% 
                          filter(income_level == "H"))
#summary(high_inc_lm)
coeftest(high_inc_lm, vcov = vcovHC(high_inc_lm, type = "HC1")) #Robust SE

# 3. Upper-middle income countries
upper_mid_lm <- lm(sdg_overall ~ spi_comp + di_score + log_gdppc + population + factor(year),
                         data = merged_2015 %>% 
                           filter(income_level == "UM"))
#summary(upper_mid_lm)
coeftest(upper_mid_lm, vcov = vcovHC(upper_mid_lm, type = "HC1")) #Robust SE

# 4. Lower-middle income countries
lower_mid_lm <- lm(sdg_overall ~ spi_comp + di_score + log_gdppc + population + factor(year),
                         data = merged_2015 %>% 
                           filter(income_level == "LM"))
#summary(lower_mid_lm)
coeftest(lower_mid_lm, vcov = vcovHC(lower_mid_lm, type = "HC1")) #Robust SE

# 5. Low income countries
low_inc_lm <- lm(sdg_overall ~ spi_comp + di_score + log_gdppc + population + factor(year),
                       data = merged_2015 %>% 
                         filter(income_level == "L"))
#summary(low_inc_lm)
coeftest(low_inc_lm, vcov = vcovHC(low_inc_lm, type = "HC1")) #Robust SE
```


### GNI Classifications DF
```{r echo=FALSE}
# 1. Overall model
overall_ct <- coeftest(overall_lm, vcov = vcovHC(overall_lm, type = "HC1"))
overall_df <- data.frame(
  model = "overall_lm",
  term = rownames(overall_ct),
  estimate = overall_ct[, "Estimate"],
  std.error = overall_ct[, "Std. Error"],
  t.statistic = overall_ct[, "t value"],
  p.value = overall_ct[, "Pr(>|t|)"],
  n = nobs(overall_lm)
)

# 2. High income countries
high_inc_ct <- coeftest(high_inc_lm, vcov = vcovHC(high_inc_lm, type = "HC1"))
high_inc_df <- data.frame(
  model = "high_inc_lm",
  term = rownames(high_inc_ct),
  estimate = high_inc_ct[, "Estimate"],
  std.error = high_inc_ct[, "Std. Error"],
  t.statistic = high_inc_ct[, "t value"],
  p.value = high_inc_ct[, "Pr(>|t|)"],
  n = nobs(high_inc_lm)
)

# 3. Upper-middle income countries
upper_mid_ct <- coeftest(upper_mid_lm, vcov = vcovHC(upper_mid_lm, type = "HC1"))
upper_mid_df <- data.frame(
  model = "upper_mid_lm",
  term = rownames(upper_mid_ct),
  estimate = upper_mid_ct[, "Estimate"],
  std.error = upper_mid_ct[, "Std. Error"],
  t.statistic = upper_mid_ct[, "t value"],
  p.value = upper_mid_ct[, "Pr(>|t|)"],
  n = nobs(upper_mid_lm)
)

# 4. Lower-middle income countries
lower_mid_ct <- coeftest(lower_mid_lm, vcov = vcovHC(lower_mid_lm, type = "HC1"))
lower_mid_df <- data.frame(
  model = "lower_mid_lm",
  term = rownames(lower_mid_ct),
  estimate = lower_mid_ct[, "Estimate"],
  std.error = lower_mid_ct[, "Std. Error"],
  t.statistic = lower_mid_ct[, "t value"],
  p.value = lower_mid_ct[, "Pr(>|t|)"],
  n = nobs(lower_mid_lm)
)

# 5. Low income countries
low_inc_ct <- coeftest(low_inc_lm, vcov = vcovHC(low_inc_lm, type = "HC1"))
low_inc_df <- data.frame(
  model = "low_inc_lm",
  term = rownames(low_inc_ct),
  estimate = low_inc_ct[, "Estimate"],
  std.error = low_inc_ct[, "Std. Error"],
  t.statistic = low_inc_ct[, "t value"],
  p.value = low_inc_ct[, "Pr(>|t|)"],
  n = nobs(low_inc_lm)
)

# Combine all results
gni_classes_ols <- bind_rows(
  overall_df,
  high_inc_df,
  upper_mid_df,
  lower_mid_df,
  low_inc_df
)

attr(gni_classes_ols $std.error, "label") <- "Robust Std. Errors Adjusted"
attr(gni_classes_ols $t.statistic, "label") <- "Robust Std. Errors Adjusted"
attr(gni_classes_ols $p.value, "label") <- "Robust Std. Errors Adjusted"

gni_classes_ols

#interactive table for side access 
#datatable(gni_classes_ols, caption = "Regression Results, by GNI Country Classifications")

# write to directory 
write.csv(gni_classes_ols, "component_1/results_csv/gni_classes_ols.csv")
```


### Visualizing Slopes: plotting multiple regression - by subgroup
```{r echo=FALSE}
viz_gni_class <- ggplot(data = merged_2015, aes(x = spi_comp, 
                                           y = sdg_overall, 
                                           color = income_level_lab)) +
  geom_point(alpha = 0.25, size = 0.75) + 
  # Overall regression line (black)
  geom_smooth(aes(group = 1), 
              method = "lm", 
              linewidth = 0.75, 
              se = FALSE, 
              color = "black") +
  # Group-specific regression lines
  geom_smooth(method = "lm", 
              linewidth = 0.65, 
              se = FALSE) +
  scale_color_manual(
    values = c("High Income Countries" = "#1D6A96", 
               "Upper-Middle Income Countries" = "#4CB5AE",
               "Lower-Middle Income Countries" = "#F3A738",
               "Low Income Countries" = "#C02942")
  ) +
  labs(title = "Relationship between SPI and SDG by World Bank Income Classification",
       x = "Statistical Performance Indicators (SPI)",
       y = "Sustainable Development Goals (SDG)",
       color = "Income Classification") +
  theme_bw() 

viz_gni_class
#ggplotly(viz_gni_class)

# Save to specific folder
ggsave("component_1/figures/gni_subgroups_ols.png", viz_gni_class, width = 10, height = 6)
```
[more results TBD]


#### Coefficient & Interval Plot
```{r echo=FALSE}
# New fd with SPI coefficients data
spi_plot_data <- data.frame(
  model = c("overall_lm", "high_inc_lm", "upper_mid_lm", "lower_mid_lm", "low_inc_lm"),
  estimate = c(0.286414727883271, 0.355563975838364, 0.174313752024353, 0.382217438772999, 0.169651614020134),
  std.error = c(0.0156271299809294, 0.0170293332032076, 0.0139294904502819, 0.0262707606446246, 0.0441323334102258)
)

# Calculate confidence intervals
spi_plot_data <- spi_plot_data %>%
  mutate(
    CI_lower = estimate - 1.96 * std.error,
    CI_upper = estimate + 1.96 * std.error
  )

# Set model order
model_order <- c("low_inc_lm", "lower_mid_lm", "upper_mid_lm", "high_inc_lm", "overall_lm")
spi_plot_data$model <- factor(spi_plot_data$model, levels = model_order)

# Create the coefficient plot
coef_inter_spi_plot <- ggplot(spi_plot_data, aes(x = estimate, y = model)) +
  geom_point(size = 3, color = "dodgerblue") +
  geom_errorbarh(aes(xmin = CI_lower, xmax = CI_upper), 
                 height = 0.2, 
                 color = "dodgerblue") +
  labs(
    title = "Coefficient Estimates with 95% Confidence Intervals for SPI by Income Group Models",
    x = "Coefficient Estimate for SPI",
    y = NULL
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_x_continuous(limits = c(0, 0.45)) +
  scale_y_discrete(labels = c("low_inc_lm" = "Low Income", 
                              "lower_mid_lm" = "Lower-Middle Income",
                              "upper_mid_lm" = "Upper-Middle Income", 
                              "high_inc_lm" = "High Income",
                              "overall_lm" = "Overall"))

coef_inter_spi_plot

ggsave("component_1/figures/coef_inter_spi_plot.png", coef_inter_spi_plot, width = 9, height = 5)
```