---
title: "adjust_outputs"
output: html_document
---

### FIRST: Libraries, Directory & Data 

```{r include=FALSE}
# set working directory 
setwd("~/Documents/GitHub/stat_casualties_study")

#load libraries 
source("packages.R")
```

### SECOND: Run function in r-script: df_years_function.R   
*[ADJUST TIME OR SKIP IF YOU'RE LOADING DATA FROM DIRECTORY]*

```{r eval=FALSE, include=FALSE}
# df_years2.0():[Updated Version] 
# Function used for extracting all available country data from datasets in specified year range

source("wrangling/df_years2.0_Function.R") #load function
merged_full <- df_years2.0(2004, 2023) #specify start & end year

# Robustness check for missing identifier variables: country_code, country_name and year
check_missing_identifiers <- function(df) {
  missing_rows <- sum(is.na(df$country_code) | is.na(df$country_name) | is.na(df$year))
  print(paste("Number of rows with missing identifiers:", missing_rows))
  return(missing_rows)}
missing_ids <- check_missing_identifiers(merged_full)

# NEW full version as CSV output (unhash when certain)
write_csv(merged_full, file = 'data/output_datasets/merged_full_NEW.csv')
```

## Cross Sectional Data
- Prioritizing SDG data
- Country-years are independent units
```{r}
# All observations with an sdg_overall value
comp1_sdg_data <- merged_full %>%
  filter(!is.na(sdg_overall)) 

#saving merged CSV output (unhash when certain)
write_csv(comp1_sdg_data, file = 'data/output_datasets/comp1_data.csv')
```


## Dependent Y = SPI (spi_comp)
**Prioritizing spi_comp data**
Deleting all countries without any SPI data 
```{r}
# load 'merged_full_NEW.csv'
merged_full <- read_csv("data/output_datasets/merged_full_NEW.csv")

# Define SDG columns
spi_cols <- c('p1_use', 'p2_services', 'p3_products', 'p4_sources', 'p5_infra')

# 1) Only countries with atleast 1 value in any of the 5 SPI pillars
all_missing_pillars <- merged_full %>% 
  group_by(country_code, country_name) %>%
  summarise(all_missing = all(if_all(all_of(spi_cols), is.na))) %>%
  filter(all_missing) %>%
  select(country_code, country_name) %>% 
  ungroup()
# Saving countries w missing SPI data
write_csv(all_missing_pillars, "wrangling/adjust_outputs_diagnostics/countries_missing_all_spi.csv")

# Removing countries from main DF 
merged_cleaned_spi_1 <- merged_full %>% 
  filter(!country_code %in% all_missing_pillars$country_code)

# 2) Only countries with atleast 1 value in the SPI composite (spi_comp) in any year
all_missing_spi_comp <- merged_cleaned_spi_1 %>%
  group_by(country_code, country_name) %>%
  summarise(all_missing = all(is.na(spi_comp))) %>%
  filter(all_missing) %>%
  select(country_code, country_name) %>% 
  ungroup()
# Saving countries w missing SPI data
write_csv(all_missing_spi_comp, "wrangling/adjust_outputs_diagnostics/countries_missing_spi_comp.csv")

# Removing countries from main DF
merged_cleaned_spi_2 <- merged_cleaned_spi_1 %>%
  filter(!country_code %in% all_missing_spi_comp$country_code)

# 3) Only countries with 70% of SPI Composite values across all years (2015-2023)

# Define threshold 
threshold <- 70 # 70% of SPI index values must be present for each country to remain

insufficient_spi_comp <- merged_cleaned_spi_2 %>%
  filter(year >= 2015) %>% # Ensure we only consider years from 2015 onwards
  group_by(country_code, country_name) %>%
  summarise(pct_available = mean(!is.na(spi_comp)) * 100) %>%
  filter(pct_available < threshold) %>%
  select(country_code, country_name) %>% 
  ungroup()
# Saving countries w insufficient SPI data  
write_csv(insufficient_spi_comp, "wrangling/adjust_outputs_diagnostics/countries_insufficient_spi_comp.csv")

# Removing countries from main DF
merged_cleaned_spi_3 <- merged_cleaned_spi_2 %>%
  filter(year >= 2015) %>% # Ensure we only consider years from 2015 onwards
  filter(!country_code %in% insufficient_spi_comp$country_code)

# Saving final dataframe 
write_csv(merged_cleaned_spi_3, "data/output_datasets/merged_cleaned_spi.csv")
```

## SPI df Diagnostics: Countries removed from SPI analyses
```{r}
# combining dataframes of filtered countries for spi
countries_not_in_spi_analysis <- bind_rows(all_missing_pillars, all_missing_spi_comp, insufficient_spi_comp) %>%
  distinct(country_code, .keep_all = TRUE) # no duplicates

# Countries removed from SPI analysis
countries_not_in_spi_analysis1 <- read_csv("wrangling/adjust_outputs_diagnostics/countries_missing_all_spi.csv") %>%
  bind_rows(read_csv("wrangling/adjust_outputs_diagnostics/countries_missing_spi_comp.csv")) %>%
  bind_rows(read_csv("wrangling/adjust_outputs_diagnostics/countries_insufficient_spi_comp.csv")) %>%
  distinct(country_code, .keep_all = TRUE) # no duplicates

# Ensuring no discrepancies between saved and active DF
rm(check_identical_spi) # remove previous instance if it exists
check_identical_spi <- test_that("Dataframes are identical", {
  expect_identical(countries_not_in_spi_analysis, countries_not_in_spi_analysis1)
})

# Save if the check_identical test passes (==TRUE)
if (check_identical_spi==TRUE) {
  write_csv(countries_not_in_spi_analysis, "wrangling/adjust_outputs_diagnostics/countries_not_in_spi_analysis.csv")
  paste("No discrepencies, countries_not_in_spi_analysis.csv saved successfully")
} 
```

## Dependent Y = SDG (sdg_overall)
**Prioritizing SDG data**
Deleting all countries without any SDG data 
```{r}
# load 'merged_full_NEW.csv'
merged_full <- read_csv("data/output_datasets/merged_full_NEW.csv")

# Define SDG columns
sdg_cols <- paste0("goal", 1:17) 

# 1) Only countries with atleast 1 value in any of the 17 SDG goals
all_missing_sdgs <- merged_full %>% 
  group_by(country_code, country_name) %>%
  summarise(all_missing = all(if_all(all_of(sdg_cols), is.na))) %>%
  filter(all_missing) %>%
  select(country_code, country_name) %>% 
  ungroup()
# Saving countries w missing SDG data
write_csv(all_missing_sdgs, "wrangling/adjust_outputs_diagnostics/countries_missing_all_sdgs.csv")

# Removing countries from main DF 
merged_cleaned_sdg_1 <- merged_full %>% 
  filter(!country_code %in% all_missing_sdgs$country_code)

# 2) Please filter to only countries with atleast 1 value in the SDG Overall Index (sdg_overall) in any year
all_missing_sdg_overall <- merged_cleaned_sdg_1 %>%
  group_by(country_code, country_name) %>%
  summarise(all_missing = all(is.na(sdg_overall))) %>%
  filter(all_missing) %>%
  select(country_code, country_name) %>% 
  ungroup()
# Saving countries w missing SDG data
write_csv(all_missing_sdg_overall, "wrangling/adjust_outputs_diagnostics/countries_missing_sdg_overall.csv")

# Removing countries from main DF
merged_cleaned_sdg_2 <- merged_cleaned_sdg_1 %>%
  filter(!country_code %in% all_missing_sdg_overall$country_code)

# 3) Only countries with 70% of SDG Index values across all years (2004-2023)
threshold <- 70 # 70% of SDG index values must be present for each country to remain
insufficient_sdg_overall <- merged_cleaned_sdg_2 %>%
  group_by(country_code, country_name) %>%
  summarise(pct_available = mean(!is.na(sdg_overall)) * 100) %>%
  filter(pct_available < threshold) %>%
  select(country_code, country_name) %>% 
  ungroup()
# Saving countries w insufficient SDG data  
write_csv(insufficient_sdg_overall, "wrangling/adjust_outputs_diagnostics/countries_insufficient_sdg_overall.csv")

# Removing countries from main DF
merged_cleaned_sdg_3 <- merged_cleaned_sdg_2 %>%
  filter(!country_code %in% insufficient_sdg_overall$country_code)

# Saving final dataframe 
write_csv(merged_cleaned_sdg_3, "data/output_datasets/merged_cleaned_sdg.csv")
```

## SDG df Diagnostics: Countries removed from SDG analysis
```{r}
# combining dataframes of filtered countries for sdg
countries_not_in_sdg_analysis <- bind_rows(all_missing_sdgs, all_missing_sdg_overall, insufficient_sdg_overall) %>%
  distinct(country_code, .keep_all = TRUE) # Ensure no duplicates

# Countries removed from SPI analysis
countries_not_in_sdg_analysis1 <-
  read_csv("wrangling/adjust_outputs_diagnostics/countries_missing_all_sdgs.csv") %>%
  bind_rows(read_csv("wrangling/adjust_outputs_diagnostics/countries_missing_sdg_overall.csv")) %>%
  bind_rows(read_csv("wrangling/adjust_outputs_diagnostics/countries_insufficient_sdg_overall.csv")) %>%
  distinct(country_code, .keep_all = TRUE) # Ensure no duplicates

# Ensuring no discrepancies between saved and active DF
check_identical_sdg <- test_that("Dataframes are identical", {
  expect_identical(countries_not_in_sdg_analysis, countries_not_in_sdg_analysis1)
})

# save if the check_identical test passes (==TRUE)
if (check_identical_sdg==TRUE) {
  write_csv(countries_not_in_sdg_analysis, "wrangling/adjust_outputs_diagnostics/countries_not_in_sdg_analysis.csv")
  paste("No discrepencies, countries_not_in_sdg_analysis.csv saved successfully")
} 

```

## Compare missing countries between SPI and SDG analyses
**Please note:** we are comparing dataframes of missing countries NOT the cleaned dataframes of countries that are used in analyses
```{r}
# Number countries not in either analysis 
overlap_countries <- intersect(countries_not_in_spi_analysis$country_code, countries_not_in_sdg_analysis$country_code)
print(paste("Number of countries removed from both SPI and SDG analyses:", length(overlap_countries)))

# Number of countries Missing in SPI that are not Missing in SDG
unique_to_spi <- setdiff(countries_not_in_spi_analysis$country_code, countries_not_in_sdg_analysis$country_code)
print(paste("Number of countries removed only from SPI analysis:", length(unique_to_spi)))

# Number of countries Missing in SDG that are not Missing in SPI
unique_to_sdg <- setdiff(countries_not_in_sdg_analysis$country_code, countries_not_in_spi_analysis$country_code)
print(paste("Number of countries removed only from SDG analysis:", length(unique_to_sdg)))

## Comparison dataframe: showing which countries were removed from which analysis
countries_removed_comparison <- bind_rows(
  countries_not_in_spi_analysis %>% mutate(removed_from = "SPI"),
  countries_not_in_sdg_analysis %>% mutate(removed_from = "SDG")
) %>%
  distinct(country_code, removed_from, .keep_all = TRUE) %>%
  pivot_wider(names_from = removed_from, values_from = removed_from, values_fill = NA) %>%
  mutate(removed_from = case_when(
    !is.na(SPI) & !is.na(SDG) ~ "Not in Either Analysis",
    !is.na(SPI) ~ "Not in SPI Analysis",
    !is.na(SDG) ~ "Not in SDG Analysis"
  )) %>%
  select(country_code, country_name, removed_from)
View(countries_removed_comparison)

# Check: save if countries listed as "Not in SPI Analysis" are not found in the SPI cleaned DF and vice versa, stop if any country listed as "Not in SPI Analysis" is found in the cleaned SPI DF and vice versa
rm(check_rm_compare) # remove previous instance if it exists
check_rm_compare <- test_that("Countries labeled as 'Not in SPI Analysis' or 'Not in SDG Analysis' are not in the respective cleaned dataframes", {
  expect_true(all(!countries_removed_comparison %>%
                    filter(removed_from == "Not in SPI Analysis") %>%
                    pull(country_code) %in% merged_cleaned_spi_3$country_code))
  expect_true(all(!countries_removed_comparison %>%
                    filter(removed_from == "Not in SDG Analysis") %>%
                    pull(country_code) %in% merged_cleaned_sdg_3$country_code))
})

# Save if the check_rm_compare test passes (==TRUE)
if (check_rm_compare==TRUE) {
  write_csv(countries_removed_comparison, "wrangling/adjust_outputs_diagnostics/all_countries_removed_comparison.csv")
  paste("No discrepencies, all_countries_removed_comparison.csv saved successfully")
} 

# Summary counts dataframe: number of countries removed from each analysis
summary_counts_all_removed_comparison <- countries_removed_comparison %>%
  group_by(removed_from) %>%
  summarise(num_countries = n()) %>%
  ungroup()
print(summary_counts_all_removed_comparison)

# saving summary_counts_all_removed_comparison
write_csv(summary_counts_all_removed_comparison, "wrangling/adjust_outputs_diagnostics/summary_counts_all_removed_comparison.csv")
```

## Create Inclusive and Exclusive DFs based on SPI and SDG analyses
```{r}
# Inclusive DF: Countries in SPI or SDG analyses
inclusive_countries <- union(merged_cleaned_spi_3$country_code, merged_cleaned_sdg_3$country_code)
merged_inclusive <- merged_full %>%
  filter(country_code %in% inclusive_countries)

# Exclusive DF: Countries in both SPI and SDG analyses
exclusive_countries <- intersect(merged_cleaned_spi_3$country_code, merged_cleaned_sdg_3$country_code)
merged_exclusive <- merged_full %>%
  filter(country_code %in% exclusive_countries)

# run check: ensures that no countries in exclusive DFs are in the missing countries list
rm(check_exclusive) # remove previous instance if it exists
check_exclusive <- test_that("No countries in exclusive DF are in missing countries list", {
  expect_true(all(!merged_exclusive$country_code %in% countries_not_in_spi_analysis$country_code))
  expect_true(all(!merged_exclusive$country_code %in% countries_not_in_sdg_analysis$country_code))
})

# run check: ensures that only countries labelled as 'Not in SPI Analysis' and 'Not in SDG Analysis' under the removed_from column in the missing countries list are still included in the inclusive DF alonhgg with countries in both analyses
rm(check_inclusive) # remove previous instance if it exists
check_inclusive <- test_that("All countries labeled as 'Not in SPI Analysis' and 'Not in SDG Analysis' are in inclusive DF", { 
  expect_true(all(countries_removed_comparison %>%
                    filter(removed_from %in% c("Not in SPI Analysis", "Not in SDG Analysis")) %>%
                    pull(country_code) %in% merged_inclusive$country_code))
})

# Saves DFs if both checks pass 
if (check_exclusive==TRUE & check_inclusive==TRUE) {
  write_csv(merged_inclusive, "data/output_datasets/merged_inclusive.csv")
  write_csv(merged_exclusive, "data/output_datasets/merged_exclusive.csv")
  paste("No discrepencies in inclusive and exclusive DFs, merged_inclusive.csv and merged_exclusive.csv saved successfully")
} 

# Dataframe based on merged_inclusive indicating which countries are in each dataframe
countries_studied_comparison <- merged_inclusive %>%
  distinct(country_code, country_name) %>%
  mutate(in_merged_cleaned_spi = if_else(country_code %in% merged_cleaned_spi_3$country_code, 1, 0),
         in_merged_cleaned_sdg = if_else(country_code %in% merged_cleaned_sdg_3$country_code, 1, 0), 
         in_merged_exclusive = if_else(country_code %in% merged_exclusive$country_code, 1, 0),
         in_merged_inclusive = 1) %>% # all countries in this DF are in inclusive DF
  select(country_code, country_name, in_merged_cleaned_spi, in_merged_cleaned_sdg, in_merged_exclusive, in_merged_inclusive)
View(countries_studied_comparison) # View list 

# saving countries_studied_comparison
write_csv(countries_studied_comparison, "wrangling/adjust_outputs_diagnostics/all_countries_studied_comparison.csv")

# Summary counts of countries included in each dataframe (SPI, SDG, Inclusive and Exclusive)
summary_counts_all_dfs <- tibble(
  dataframe = c("SPI Analysis", "SDG Analysis", "Inclusive DF", "Exclusive DF"),
  num_countries = c(
    sum(countries_studied_comparison$in_merged_cleaned_spi==1),
    sum(countries_studied_comparison$in_merged_cleaned_sdg==1),
    sum(countries_studied_comparison$in_merged_inclusive==1),
    sum(countries_studied_comparison$in_merged_exclusive==1)
  )
)
print(summary_counts_all_dfs)
# saving summary_counts_all_dfs
write_csv(summary_counts_all_dfs, "wrangling/adjust_outputs_diagnostics/summary_counts_all_studied_comparison.csv")
```

## Check: Number of countries with spi_comp vs sdg_overall in original merged_full dataframe for robustness
```{r}
# checking the number of countries covered by spi_comp vs sdg_overall variables in merged_full dataframe
merged_full_sub <- read_csv("data/output_datasets/merged_full_NEW.csv")
num_countries_spi <- merged_full_sub %>%
  filter(!is.na(spi_comp)) %>%
  summarise(num_countries = n_distinct(country_code)) %>%
  pull(num_countries)
num_countries_sdg <- merged_full %>%
  filter(!is.na(sdg_overall)) %>%
  summarise(num_countries = n_distinct(country_code)) %>%
  pull(num_countries)
print(paste("Number of countries with spi_comp data:", num_countries_spi))
print(paste("Number of countries with sdg_overall data:", num_countries_sdg))
```
- merged_full (base dataframe) = 187 countries with spi_comp data and 167 countries with sdg_overall data

**As of September 1 2025:**
- SPI Analysis = 174 countries (merged_cleaned_spi.csv)
- SDG Analysis = 167 countries (merged_cleaned_sdg.csv)
- Inclusive DF = 179 countries (merged_inclusive.csv)
- *Exclusive DF = 162 countries (merged_exclusive.csv)* [MAIN DATAFRAME USED IN ANALYSES]

```{r}
# saving main DF as rds for faster loading
saveRDS(merged_exclusive, file = "data/Main RDS Outputs/merged_exclusive.rds")
# To load the rds file in future analyses
# merged_exclusive <- readRDS("data/Main RDS Outputs/merged_exclusive.rds")
```


# Test for identical DFs
As needed... 
```{r}
library(testthat)

# In a test script
test_that("Dataframes are identical", {
  expect_identical(countries_not_in_spi_analysis, countries_not_in_spi_analysis1)
})
```

