---
title: "backsliding_countries"
output: html_document
---

# preparation 
```{r}
setwd("~/Documents/GitHub/QMSS_Thesis_Sanchez")

#load libraries 
source("packages.R")

#load data 
source("Comp2_panel_wrangling.R")

#selecting vars
event_data <- panel_data %>% 
  dplyr::select(country_code, year, missingness, elect_dem, aut_ep, dem_ep, regch_event, regime_type_2, regime_type_4, lib_dem, part_dem, delib_dem, egal_dem, academ_free, income_level, income_level_lab, log_gdppc, population) %>% 
  arrange(country_code, year) %>%  # Critical for correct lagging
  filter(year >= 2016)
```

# country_events dataframe: identifies/summarizes transition events per country
```{r}
#counting number of event-years: dem_ep & aut_ep
country_events <- event_data %>%
  group_by(country_code) %>%
  summarize(
    aut_ep_duration = sum(aut_ep == 1, na.rm = TRUE),
    dem_ep_duration = sum(dem_ep == 1, na.rm = TRUE), 
  )

# creating condition dummy variables
country_events <- country_events %>%
  mutate(
    has_aut_ep = (aut_ep_duration > 0) * 1, # experienced atleast 1 aut episode 
    has_dem_ep = (dem_ep_duration > 0) * 1, # experienced atleast 1 dem episode 
    either_event = (aut_ep_duration > 0 | dem_ep_duration > 0) * 1, # experience either kind of event
    neither_event = (aut_ep_duration == 0 & dem_ep_duration == 0) * 1 # experienced neither event 
  )

# frequency summary stats [inclusive]
summary_counts_incl <- country_events %>%
  summarize(
    num_countries_with_aut_events = sum(has_aut_ep),
    num_countries_with_dem_events = sum(has_dem_ep),
    num_countries_with_events = sum(either_event),
    num_countries_without_events = sum(neither_event), 
    total_countries = n()
  )

# labels for exclusive sorting [exclusive]
country_events <- country_events %>%
  mutate(event_pattern_excl = case_when(
    has_aut_ep == 1 & has_dem_ep == 1 ~ "Both Events", #must go first to label both events 
    has_aut_ep == 1 ~ "Backsliding Only",
    has_dem_ep == 1 ~ "Democratizing Only",
    neither_event == 1 ~ "Neither Event",
  ))

# frequency summary stats [exclusive]
summary_counts_excl <- country_events %>%
  summarize(
    num_countries_with_only_aut_events = sum(event_pattern_excl== "Backsliding Only"),
    num_countries_with_only_dem_events = sum(event_pattern_excl== "Democratizing Only"),
    num_countries_with_both_events = sum(event_pattern_excl== "Both Events"),
    num_countries_with_neither_events = sum(event_pattern_excl== "Neither Event"),
    total_countries = n()
  )
```

# creating summary table of event patterns (CSV & Stargazer Table)
```{r}
# Frequency & proportion table of event patterns [inclusive & exclusive]
country_events_summary <- data.frame(
  Event_Pattern = c("Countries with Autocratization Events", 
                    "Countries with Democratization Events", 
                    "Countries with Either Event", 
                    "Countries with ONLY Autocratization Events",
                    "Countries with ONLY Democratization Events",
                    "Countries with Both Events",
                    "Countries with Neither Event",
                    "Total Countries"),
  Frequency = c(summary_counts_incl$num_countries_with_aut_events, 
                summary_counts_incl$num_countries_with_dem_events, 
                summary_counts_incl$num_countries_with_events, 
                summary_counts_excl$num_countries_with_only_aut_events,
                summary_counts_excl$num_countries_with_only_dem_events,
                summary_counts_excl$num_countries_with_both_events,
                summary_counts_excl$num_countries_with_neither_events,
                summary_counts_excl$total_countries
                ),
  Proportion = c(summary_counts_incl$num_countries_with_aut_events / summary_counts_incl$total_countries,
                 summary_counts_incl$num_countries_with_dem_events / summary_counts_incl$total_countries,
                 summary_counts_incl$num_countries_with_events / summary_counts_incl$total_countries,
                 summary_counts_excl$num_countries_with_only_aut_events / summary_counts_excl$total_countries,
                 summary_counts_excl$num_countries_with_only_dem_events / summary_counts_excl$total_countries,
                 summary_counts_excl$num_countries_with_both_events / summary_counts_excl$total_countries,
                 summary_counts_excl$num_countries_with_neither_events / summary_counts_excl$total_countries,
  summary_counts_excl$total_countries / summary_counts_excl$total_countries
))

# write to csv
#write_csv(country_events_summary, "results_csv/country_events_summary.csv")

# creating stargazer table object
country_events_table <- stargazer(country_events_summary, 
                                    summary = FALSE, 
                                    title = "Summary of Transition Event Patterns in 167 Countries (2015-2023)", 
                                    type = "text", 
                                    digits = 2,
                                    rownames = FALSE,
                                    align = TRUE,
                                    no.space = TRUE,
                                    font.size = "small",
                                    #style = "ajps",
                                    label = "tab:country_events_summary", 
                                    out = "figures/country_events_summary.html"
                                  )
# print table 
print(country_events_table)
```

**Frequency [inclusive]** 
167: countries total; 
within the time period of 2015-2023:  

- 60 countries experienced at least 1 episode of autocratization  - 35 countries experienced at least 1 episode of democratization  
- 75 countries experienced at least 1 episode of either autocratization or democratization  
- 92 countries experienced neither kind of episode, autocratization or democratization

**Proportions [inclusive]**
- The proportion of countries that have experienced at least 1 episode of autocratization in the time period between 2015 and 2023 is 60/167 = 0.3592814 (35.93%)
- The proportion of countries that have experienced at least 1 episode of democratization in the time period between 2015 and 2023 is 35/167 = 0.2095808 (20.96%)
- The proportion of countries that have experienced either kind of transition episode (democratization or autocratization) in the time period between 2015 and 2023 is 75/167 =  0.4491018 (44.91%)
- The proportion of countries that have NOT experienced any kind of transition episode the time period between 2015 and 2023 is 92/167 = 0.5508982 (55.09%)

**Frequency [exclusive]** 
- 40 countries experienced ONLY autocratization events
- 15 countries experienced ONLY democratization events
- 20 countries experienced both events (democratization and autocratization)

**Proportions [exclusive]**
- The proportion of countries that have experienced both events (democratization and autocratization) in the time period between 2015 and 2023 is 20/167 = 0.11976 (11.98%)
- The proportion of countries that have experienced only backsliding in the time period between 2015 and 2023 is 40/167 = 0.23952 (23.95%)
- The proportion of countries that have experienced only democratization in the time period between 2015 and 2023 is 15/167 = 0.08982 (8.98%)
- The proportion of countries that have experienced neither event in the time period between 2015 and 2023 is 92/167 = 0.55090 (55.09%)

# merging country_events df with event_data
```{r}
# Adding country_events data to event_data
event_data <- event_data %>%
  left_join(country_events, by = 'country_code')
  #filter(has_aut_ep == TRUE | has_dem_ep == TRUE) #FILTERS TO ONLY CHANGING REGIMES

# subgroups: regressing and democratizing coutries 
aut_data <- event_data %>% filter(has_aut_ep == 1) ## experienced at least 1 backsliding event 
dem_data <- event_data %>% filter(has_dem_ep == 1) ## experienced at least 1 democratizing event 
```

# complete regime change
```{r}
# Frequency/proportion table: instances of backsliding, stability, democratization from 2004-2024
table(event_data$regch_event) # frequency
prop.table(table(event_data$regch_event)) # proportions 

#Frequency of countries that experienced democratization 2015-2024
hist(dem_data$year[dem_data$regch_event==1], breaks = 20)

#Frequency of countries that experienced autocratization 2015-2024
hist(aut_data$year[aut_data$regch_event==-1], breaks = 20)

#Frequency of countries that experienced no change 2015-2024 [FIX HERE AND BELOW]
hist(merged$year[merged$regch_event==0], breaks = 20)
```


# Backsliding FD Models 

## Stage 1
```{r}
# SPI ~ DI (M ~ X) 
fd_spi_di_aut <- plm(
  formula = spi_comp ~ dplyr::lag(di_score, 1) + log_gdppc,
  index = c("country_code", "year"),
  data = aut_data, 
  model = "fd" #FD
) 
# Country-level custard Robust SEs
summary(fd_spi_di_aut, vcov = vcovHC(fd_spi_di_aut, cluster = "group", type = "HC1"))
```

## Stage 2
```{r}
# SDG ~ SPI (Y ~ M)
fd_sdg_spi_aut <- plm(
  formula = sdg_overall ~ lag(spi_comp, 1) + dplyr::lag(di_score, 2) + log_gdppc,
  index = c("country_code", "year"),
  data = aut_data, 
  model = "fd"
)
# Country-level custard Robust SEs
summary(fd_sdg_spi_aut, vcov = vcovHC(fd_sdg_spi_aut, cluster = "group", type = "HC1"))
```
On average, when backsliding countries experience a change in Democracy score, is there a change in statistical capacity (SPI)? On average, when backsliding countries experience a change in statistical capacity (SPI) do they similalry experience a change in SDG score? 

# Backsliding FE Models
```{r}
###### SPI ~ DI (M ~ X) ######
fe_spi_di_aut <- plm(
  formula = spi_comp ~ di_score + log_gdppc + factor(year),
  index = c("country_code", "year"),
  data = aut_data,
  model = "within" 
)
summary(fe_spi_di_aut, vcov = vcovHC(fe_spi_di_aut, cluster = "group", type = "HC1")) # Robust SEs

###### SDG ~ SPI (Y ~ M) ######
fe_sdg_spi_aut <- plm(
  sdg_overall ~ spi_comp + di_score + log_gdppc + factor(year),
  index = c("country_code", "year"),
  data = aut_data,
  model = "within" 
)
summary(fe_sdg_spi_aut, vcov = vcovHC(fe_sdg_spi_aut, cluster = "group", type = "HC1")) # Robust SEs
```

# Single Country: Hungary 
```{r}
#filter for hungary 
hun_regressed <- aut_data %>% 
  filter(country_code == "HUN")

# FE model - Hungary 
fe_testing <- plm(
  formula = spi_comp ~ di_score + log_gdppc,
  data = aut_data, sub = (country_code == 'HUN'),
  index = c("country_code", "year"),
  model = "within")
summary(fe_testing, vcov = vcovHC(fe_testing, type = "HC3"))
```

#testing for autocorrelation 
```{r}
lmtest::bgtest(fe_testing)
lmtest::bgtest(fe_testing, order = 1)
```


```{r}
# subset of countries that experienced atleast 1 democratization event 
democratized <- merged %>% 
  group_by(country_code) %>%
  filter(dem_ep == 1) %>%
  ungroup()
```
