---
title: "H2_SPI~DI"
output: html_document
---


# Set up and Wrangling 
```{r}
# set working directory
setwd("~/Documents/GitHub/QMSS_Thesis_Sanchez")

#load libraries/packages
source("packages.R")

# load data
source("Comp2_panel_wrangling.R") 
# path = 'data/Main CSV Outputs/merged_cleaned_spi.csv'
```

# Model Observations Function
Function that extracts model data for plm models
```{r}
# load the extract_plm_data function
source("functions/extract_plm_data_function.R")
```

# SPI and DI Relationship
## Baseline Model [FOUND SOMETHING!!!]
```{r}
# baseline model
FE_all <- plm(formula = spi_comp ~ di_score + plm::lag(di_score, 1) + plm::lag(di_score, 2) + plm::lag(di_score, 3) + log_gdppc + income_level_recoded,
              model = "within", 
              index = c("country_code", "year"),
              data = panel_data, 
              effect = "twoways")
summary(FE_all)
summary(FE_all, vcov = vcovHC(FE_all, cluster = "group", type = "HC1"))

# observations used
FE_all_data <- extract_plm_data(FE_all)
```


## Stable Regime, Subgrouping [NOTHING]
Using fixed effects regression, how does a change in DI score impact SPI performance? Is the impact lagged or immediate?
```{r}
# baseline model
FE_stable <- plm(formula = spi_comp ~ di_score + plm::lag(di_score, 1) + plm::lag(di_score, 2) + log_gdppc + income_level_recoded,
              model = "within", 
              index = c("country_code", "year"),
              subset = eiu_stable == 1,
              data = panel_data, 
              effect = "twoways")
summary(FE_stable)
summary(FE_stable, vcov = vcovHC(FE_all, cluster = "group", type = "HC1"))

# observations used 
FE_stable <- extract_plm_data(FE_stable)
```


## Complete Regime Change (autocratized), Subgrouping 
Of countries that experienced complete change from democracy to autocracy (15 countries), how did the decrease in DI score impact their SPI performance? Was the impact lagged or immediate?
```{r}
# autocratized variable 
FE_aut <- plm(formula = spi_comp ~ di_score + plm::lag(di_score, 1) + plm::lag(di_score, 2) + log_gdppc + income_level_recoded,
              model = "within", 
              index = c("country_code", "year"),
              subset = eiu_autocratized == 1, # only autocratized countries
              data = panel_data, 
              effect = "twoways")
summary(FE_aut)
summary(FE_aut, vcov = vcovHC(FE_aut, cluster = "group", type = "HC1"))

# observations used 
FE_aut_data <- extract_plm_data(FE_aut)
```

## Complete Regime Change (democratized), Subgrouping 
Of countries (15 countries) that experienced complete change from autocracy to democracy, how did the decrease in DI score impact their SPI performance? Was the impact lagged or immediate?
```{r}
# democratized variable 
FE_dem <- plm(formula = spi_comp ~ di_score + plm::lag(di_score, 1) + plm::lag(di_score, 2) + log_gdppc + income_level_recoded,
              model = "within", 
              index = c("country_code", "year"),
              data = panel_data, 
              subset = eiu_democratized == 1, # modifying based on regime stability
              effect = "twoways")
summary(FE_dem)
summary(FE_dem, vcov = vcovHC(FE_dem, cluster = "group", type = "HC1"))

# observations used 
FE_dem_data <- extract_plm_data(FE_dem)
```

## Put all three models in a table
```{r}
# Extracting robust standard errors for the three FE models
rob_se_fe <- list(
  coeftest(FE_all, vcov = vcovHC(FE_all, cluster = "group", type = "HC1"))[, 2],
  coeftest(FE_aut, vcov = vcovHC(FE_aut, cluster = "group", type = "HC1"))[, 2],
  coeftest(FE_dem, vcov = vcovHC(FE_dem, cluster = "group", type = "HC1"))[, 2]
)
# stargazer table for FE models
stargazer(
  FE_all, FE_aut, FE_dem,
  se = rob_se_fe,
  type = "html",
  title = "H4: SPI and DI Fixed Effects Models (Regime Change)",
  align = TRUE,
  dep.var.labels = "SPI Overall Score",
  column.labels = c("All", "Autocratized", "Democratized"),
  covariate.labels = c("DI Composite", "Lagged DI (t-1)", "Lagged DI (t-2)",  "Log(GDPper-capita)", "Lower-Middle Income (LMIC)", "Upper-Middle Income (UMIC)", "High Income (HIC)", "Intercept"),
  report = "vcs*",
  notes = "Robust standard errors clustered by country in parentheses",
  omit = "factor\\(year\\)",
  model.names = FALSE,
  out = "compiled_materials/results/h4_spi_di_FE_models.html"
)

```

# Testing other things ---------------------------------------------------------

### Using di_score_lag1 and di_score_lag2 variables (eiu_autocratized == 1)
```{r}
# autocratized variable 
FE_test <- plm(formula = spi_comp ~ di_score + di_score_lag1 + di_score_lag2 + log_gdppc + income_level_recoded,
              model = "within", 
              index = c("country_code", "year"),
              subset = eiu_autocratized == 1, # only autocratized countries
              data = panel_data, 
              effect = "twoways")
summary(FE_test)
summary(FE_test, vcov = vcovHC(FE_test, cluster = "group", type = "HC1"))
```
[DIFFERENT RESULTS, INVESTIGATE!!]

### Using di_score_lag1 and di_score_lag2 variables (eiu_democratized == 1)
```{r}
FE_test <- plm(formula = spi_comp ~ di_score + di_score_lag1 + di_score_lag2 + log_gdppc + income_level_recoded,
              model = "within", 
              index = c("country_code", "year"),
              subset = eiu_democratized == 1, # only autocratized countries
              data = panel_data, 
              effect = "twoways")
summary(FE_test)
summary(FE_test, vcov = vcovHC(FE_test, cluster = "group", type = "HC1"))
```
[DIFFERENT RESULTS, INVESTIGATE!!]

## interaction terms with plm lagged di_score
```{r}
FE_test <- plm(formula = spi_comp ~ di_score*plm::lag(di_score, 1) + plm::lag(di_score, 2) + log_gdppc + income_level_recoded,
              model = "within", 
              index = c("country_code", "year"),
              subset = eiu_autocratized == 1, # only autocratized countries
              data = panel_data, 
              effect = "twoways")
summary(FE_test)
summary(FE_test, vcov = vcovHC(FE_test, cluster = "group", type = "HC1"))

# observations used
FE_test_data <- extract_plm_data(FE_test)
```

## adjusting x inputs 
```{r}
FE_test <- plm(formula = spi_comp ~ di_score + plm::lag(di_score, 1) + plm::lag(di_score, 2) + log_gdppc + income_level_recoded,
              model = "within", 
              index = c("country_code", "year"),
              subset = eiu_stable == 1, # only autocratized countries
              data = panel_data, 
              effect = "twoways")
summary(FE_test)
summary(FE_test, vcov = vcovHC(FE_test, cluster = "group", type = "HC1"))

# observations used
FE_test_data <- extract_plm_data(FE_test)
```
